Equations & R Code

### 2016-04-27 manuscript/equations.Rmd

```{r, eval=F, echo=TRUE}
## Continuous model
updateDNR <- function(t, DNR, parms){
  with(as.list(c(DNR, parms)), {
    dD1dt = 0
    dD2dt = 0
    dN1dt = N1*eps[1]*(uptake_R(r[1], R, alpha[1], beta[1]))
    dN2dt = N2*eps[2]*(uptake_R(r[2], R, alpha[2], beta[2]))
    dRdt = -1 * (dN1dt/eps[1] + dN2dt/eps[2])
    list(c(dD1dt, dD2dt, dN1dt, dN2dt, dRdt)) #output
  })
}

##  Resource uptake function (Hill function)
uptake_R <- function(r, R, alpha, beta){
  return((r*R^alpha) / (beta^alpha + R^alpha))
}
```

@@

```{r, eval=F, echo=T}
## Discrete model
update_DNR <- function(t,DNR,gs,a1,a2,m1,m2,b1,b2,theta1,theta2,c){
  with (as.list(DNR),{
    g1 <- gs[1]
    g2 <- gs[2]
    D1 <- a1*N1 + D1*(1-g1)*(1-m1)
    D2 <- a2*N2 + D2*(1-g2)*(1-m2)
    N1 <- b1*(1-a1)*N1 + g1*(D1+(a1*N1))*(1-m1)
    N2 <- b2*(1-a2)*N2 + g2*(D2+(a2*N2))*(1-m2)
    R <- theta1*(1-a1)*N1 + theta2*(1-a2)*N2 + c*R + Rvector[t]
    return(c(D1, D2, N1, N2, R))
  })
}
```

@@

####
#### Model functions -------------------------------------------------
####
## Continuous model
updateDNR <- function(t, DNR, parms){
  with(as.list(c(DNR, parms)), {
    dD1dt = 0
    dD2dt = 0
    dN1dt = N1*eps[1]*(uptake_R(r[1], R, a[1], b[1]))
    dN2dt = N2*eps[2]*(uptake_R(r[2], R, a[2], b[2]))
    dRdt = -1 * (dN1dt/eps[1] + dN2dt/eps[2])
    list(c(dD1dt, dD2dt, dN1dt, dN2dt, dRdt)) #output
  })
}

@@

# Loop over seasons
nms <- names(DNR)
gVec <- getG(sigE = sigE, rho = rho, nTime = seasons)
Rvector <- rlnorm(seasons, Rmu, Rsd_annual)
saved_outs <- matrix(ncol=5, nrow=seasons)
for(season_now in 1:seasons){
  output <- ode(y = DNR, times=days,
                func = updateDNR, parms = parms)
  DNR <- output[nrow(output),nms]
  saved_outs[season_now,] <- DNR
  names(DNR) <- nms
  DNR <- update_DNR(season_now, DNR, gVec[season_now,],
                    alpha1=alpha1,alpha2=alpha2,
                    eta1=eta1,eta2=eta2,
                    beta1=beta1,beta2=beta2,
                    theta1=theta1,theta2=theta2,nu=nu)
  names(DNR) <- nms
}
matplot(saved_outs[101:seasons,c(3:4)], type="l", 
        xlab="Season", ylab="Abundance",
        main=expression(paste(alpha[1],"= 0.5; ", alpha[2],"=0.49; ", sigma[E],"=0.2")))
```

>>>>

```{r, eval=F, echo=TRUE}
## Continuous model
updateNR <- function(t, NR, parms){
  with(as.list(c(NR, parms)), {
    dN1dt = N1*eps[1]*(uptake_R(r[1], R, a[1], b[1]))
    dN2dt = N2*eps[2]*(uptake_R(r[2], R, a[2], b[2]))
    dRdt = -1 * (dN1dt/eps[1] + dN2dt/eps[2])
    list(c(dN1dt, dN2dt, dRdt)) #output
  })
}

##  Resource uptake function (Hill function)
uptake_R <- function(r, R, a, b){
  return((r*R^a) / (b^a + R^a))
}
```

@@

```{r, eval=F, echo=T}
## Discrete model
update_DNR <- function(t,DNR,gammas,alpha1,alpha2,eta1,eta2,beta1,beta2,theta1,theta2,nu){
  with (as.list(DNR),{
    g1 <- gammas[1]
    g2 <- gammas[2]
    D1 <- alpha1*N1 + D1*(1-g1)*(1-eta1)
    D2 <- alpha2*N2 + D2*(1-g2)*(1-eta2)
    N1 <- beta1*(1-alpha1)*N1 + g1*(D1+(alpha1*N1))*(1-eta1)
    N2 <- beta2*(1-alpha2)*N2 + g2*(D2+(alpha2*N2))*(1-eta2)
    R <- theta1*(1-alpha1)*N1 + theta2*(1-alpha2)*N2 + nu*R + Rvector[t]
    return(c(D1, D2, N1, N2, R))
  })
}
```

@@

####
#### Model functions -------------------------------------------------
####
## Continuous model
updateNR <- function(t, NR, parms){
  with(as.list(c(NR, parms)), {
    dN1dt = N1*eps[1]*(uptake_R(r[1], R, a[1], b[1]))
    dN2dt = N2*eps[2]*(uptake_R(r[2], R, a[2], b[2]))
    dRdt = -1 * (dN1dt/eps[1] + dN2dt/eps[2])
    list(c(dN1dt, dN2dt, dRdt)) #output
  })
}

@@

# Loop over seasons
nmsDNR <- names(DNR)
NR <- DNR[c("N1", "N2", "R")] 
nmsNR <- names(NR)
gVec <- getG(sigE = sigE, rho = rho, nTime = seasons)
Rvector <- rlnorm(seasons, Rmu, Rsd_annual)
saved_outs <- matrix(ncol=5, nrow=seasons)
for(season_now in 1:seasons){
  output <- ode(y = NR, times=days,
                func = updateNR, parms = parms)
  NR <- output[nrow(output),nmsNR]
  DNR <- c(DNR[c("D1","D2")], NR)
  saved_outs[season_now,] <- DNR
  names(DNR) <- nmsDNR
  DNR <- update_DNR(season_now, DNR, gVec[season_now,],
                    alpha1=alpha1,alpha2=alpha2,
                    eta1=eta1,eta2=eta2,
                    beta1=beta1,beta2=beta2,
                    theta1=theta1,theta2=theta2,nu=nu)
  names(DNR) <- nmsDNR
  NR <- DNR[c("N1", "N2", "R")] 
  names(NR) <- nmsNR
}
matplot(saved_outs[101:seasons,c(3:4)], type="l", 
        xlab="Season", ylab="Abundance",
        main=expression(paste(alpha[1],"= 0.5; ", alpha[2],"=0.49; ", sigma[E],"=0.2")))
```
